---
title: 'Pun_Vincent1'
output:
  html_document: default
---

```{r setup, include = FALSE}
# DO NOT ADD OR REVISE CODE HERE
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)

```

-----

### Test Items starts from here - There are 5 sections - 50 points total ##########################

Read each question carefully and address each element.  Do not output contents of vectors or data frames unless requested.

##### Section 1: (8 points) This problem deals with vector manipulations.

(1)(a)  Create a vector that contains the following, in this order, and output the final, resulting vector.  Do not round any values, unless requested.
* A sequence of integers from 0 to 4, inclusive.
* The number 13
* Three repetitions of the vector c(2, -5.1, -23).
* The arithmetic sum of 7/42, 3 and 35/42

```{r test1a}
#seq, integer, rep, c 
problem1a <-c(seq(from = 0, to = 4, by = 1),13,rep(c(2,-5.1,-23),3),((7/42)+3+(35/42)))


```

(1)(b)  Sort the vector created in (1)(a) in ascending order.  
Output this result. 
Determine the length of the resulting vector and assign to "L". 
Output L. 
Generate a descending sequence starting with L and ending with 1. 
Add this descending sequence arithmetically the sorted vector. 
This is vector addition, not vector combination. 

Output the contents. 

Do not round any values.

```{r test1b}
sort(problem1a, decreasing = FALSE) #Sort the vector created in 1a in ascending order
#length output as "L" 
L <- length(sort(problem1a, decreasing = FALSE))
L #length of sorted problem1a
sort(seq(from = L, to = 1), decreasing = TRUE) #descending from L to 1

#add descending sequence arithmetically [to] the sorted vector
problem1b_sortedvector <- sort(problem1a, decreasing = FALSE) #variables for organization
problem1b_Lto1 <- sort(seq(from = L, to = 1), decreasing = TRUE) 
#vector addition
problem1b_vectoraddition <- problem1b_Lto1 + problem1b_sortedvector
#output vector addition
problem1b_vectoraddition



```

(1)(c) Extract the first and last elements of the vector you have created in (1)(b) 
to form another vector of the extracted elements. 
Form a third vector from the elements not extracted. 
Output these vectors.

```{r test1c}
#extract first and last elements of vector
problem1c_extract <- c(head(problem1b_vectoraddition, 1), #1 placement in vector
               tail(problem1b_vectoraddition, 1)) #-1 placement in vector
#not extracted
problem1c_notextracted <- c(problem1b_vectoraddition[2:15])

#output these vectors
problem1c_extract
problem1c_notextracted

```

(1)(d) Use the vectors from (c) to 
reconstruct the vector in (b). 
Output this vector. 
Sum the elements and round to two decimal places.

```{r test1d}
#VECTOR FROM (c)
#problem1c_extract
#problem1c_notextracted

#VECTOR IN (B)
#problem1b_vectoraddition

#RECONSTRUCT = put the two vectors from (c) back together so that it is the same length as the vector in (b)
problem1d_reconstructedvector <- c(problem1c_extract,problem1c_notextracted)
#output reconstructed vector
problem1d_reconstructedvector 
#round 2 decimal places c-style string formatting 
#because round() to 2 places only shows the first decimal due to 0 for second decimal
sprintf("%.2f", sum(problem1d_reconstructedvector))

```

-----

##### Section 2: (10 points) The expression y = sin(x/2) + cos(x/2) is a trigonometric function.

(2)(a) Create a user-defined function - via *function()* - 
that implements the trigonometric function above, 
accepts numeric values, "x," 
calculates and returns values "y."

```{r test2a}
p2a_function <- function(x){
  y <- sin(x/2) + cos(x/2) #trig function
  result <- return(y)
  result #returns y 
}
```

(2)(b) Create a vector, x, of 4001 equally-spaced values from -2 to 2, inclusive. 
Compute values for y using the vector x and your function from (2)(a).  
**Do not output x or y.**  
Find the value in the vector x that corresponds to the maximum value in the vector y. 
Restrict attention to only the values of x and y you have computed; 
i.e. do not interpolate. 
Round to 3 decimal places and output both the maximum y and corresponding x value.

Finding the two desired values can be accomplished in 
as few as two lines of code.  

Do not use packages or programs you may find on the internet or elsewhere. 
Do not output the other elements of the vectors x and y. 
Relevant coding methods are given in the *Quick Start Guide for R*.

```{r test2b}
?seq()
x <- seq(from = -2, to = 2, length.out = 4001) #equally spaced between -2 and 2 4001 times 
?c()

#the two things I want to value match
match1 <- round(max(p2a_function(x)),3)
match2 <- round(x,3)

#function to match values 
p2b_intersection <- function(x, y) y[match(x, y, nomatch = 0)]
p2b_intersection(match1,match2)
```

(2)(c) Plot y versus x in color, 
with x on the horizontal axis. 
Show the location of the maximum value of y determined in 2(b).  
Show the values of x and y corresponding to the maximum value of y in the display. 
Add a title and other features such as text annotations. 
Text annotations may be added via *text()* for base R plots and 
*geom_text()* or 
*geom_label()* for ggplots.

```{r test2c}

plot(x,p2a_function(x), #x values for x axis, y values is sin(x/2) + cos(x/2)
     type = "l",col = "lightgrey",xlab = "x",ylab = "y", 
     main = "Max of sin(x/2)+cos(x/2) between -2 and 2")
points(x = p2b_intersection(match1,match2), y = p2b_intersection(match1,match2), 
       col = "red", pch =20)

```

-----

##### Section 3: (8 points) This problem requires finding the point of intersection of two functions.  Using the function ```y = cos(x/2) * sin(x/2)```, 
find where the curved line y = -(x/2)^3 intersects it within the range of values used in part (2) 
(i.e. 4001 equally-spaced values from -2 to 2).  
Plot both functions on the same display, 
and show the point of intersection.  
Present the coordinates of this point as text in the display.    

```{r test3}
#library(ggplot2) 
#Wanted to use this but unsure how
#ggplot(data, aes(x=x)) + geom_line(aes(y=p3_1(x))) NOTE/REFERENCE

p3_1 <- function(x){
  y <- cos(x/2) * sin(x/2)
  result <- return(y)
  result #returns y 
}

p3_2 <- function(x){
  y <- -(x/2)^3 
  result <- return(y)
  result #returns y
}

coordx <- as.character(p3_1(x)[which(p3_1(x) == p3_2(x))])
coordy <- as.character(p3_2(x)[which(p3_1(x) == p3_2(x))])


plot(x, p3_1(x), type = "p", col = "lightgrey",
     xlab = "(x/2)^3",ylab = "cos(x/2 *  sin(x/2)")
lines(x, p3_2(x), type = "p", col = "orange")
text(x = (p3_1(x)[which(p3_1(x) == p3_2(x))]), 
     y = (p3_2(x)[which(p3_1(x) == p3_2(x))]),
     labels = expression("0,0"),
     col="black", 
     pos = 3,
     cex = 0.8)

```

```{r test3 REVISED}

x <- seq(-2, 2, length = 4001) #sequence between -2 and 2 

func1 <- cos(x/2) * sin(x/2) #function 1  

func2 <- -(x/2)^3 #function 2 

which(func1 == func2) #find where they equal eqch other 

x[which(func1 == func2)] #within x, find where func1 equals func2 

func1[which(func1 == func2)] #coordinate of func1 where func1 equals func2 

func2[which(func1 == func2)] #coordinate of func2 where func1 equsl func2 

plot(x, func1, main = "Plot of Functions", col = "red", #red func1 

     xlab = " x values",

     ylab = "y values",

     type = "l", #lines 

     xlim = c(-2.5, 2.5)) #since x is a sequuence between -2  and 2 

points(func1[which(func1 == func2)], #points where intersect  

       func2[which(func1 == func2)], #points where rintersect 

       col = rgb(0,0, 1,0.9), #color

       pch = 16,

       cex = 3)

?points()
?lines()

lines(x, func2, col = "green")

text(func1[which(func1 == func2)] - 1.5,

     func2[which(func1 == func2)] - 0.1,

     "Intersection at Origin") #HOW SHE PUT TEXT ON TOP OF THE COORDINATES 

abline(h = 0, lty = 3) #how she got the line to show at the cross section
?abline()

abline(v = 0, lty = 3)

text(x = -1.5, y = -0.2, paste("x = ", #IMPORTANT HOW SHE PUT THE TEXT + COORDINATES 

                               func1[which(func1 == func2)],

                               "  y =  ",

                               func2[which(func1 == func2)]))

 
```


-----

##### Section 4: (12 points) Use the "trees" dataset for the following items. This dataset has three variables (Girth, Height, Volume) on 31 felled black cherry trees.

(4)(a) Use *data(trees)* to load the dataset. 
Check and output the structure with *str()*. 
Use *apply()* to return the median values for the three variables. 
Output these values.  
Using R and logicals, output the row number and the three measurements - 
Girth, Height and Volume - 
of any trees with Girth equal to median Girth.  
It is possible to accomplish this last request with one line of code.  

```{r test3a}

data(trees)
#head(trees)
#?median()
apply(trees,2,median)

trees[which(trees$Girth == median(trees$Girth)),]
```

(4)(b) Girth is defined as the diameter of a tree taken at 4 feet 6 inches from the ground. 
Convert each diameter to a radius, r. 
Calculate the cross-sectional area of each tree using pi times the squared radius. 
Present a stem-and-leaf plot of the radii, 
and a histogram of the radii in color. 
Plot Area (y-axis) versus Radius (x-axis)  in color showing the individual data points.  
Label appropriately.

```{r test3b}
#radius
p4b_radius <- trees$Girth / 2
#area
p4b_area <- pi*p4b_radius**2
#stem and leaf plot of the radii
stem(p4b_radius)
#mfcol
par(mfcol = c(1,2))
#histogram of radii in color
?hist()
hist(p4b_radius, breaks = 6, 
     col = c("cadetblue1","cadetblue2","cadetblue3","cadetblue4","aquamarine","aquamarine1","aquamarine2"), 
     main ="Radii in Color",
     xlab = "Stems",
     ylab = "Leaves")
#plot area (y) versus radius (x) in color
plot(p4b_radius,#x axis
     p4b_area, #y axis
     main = "Area versus Radius",
     col = "steelblue1", #in color
     xlab = "Radius", #label x
     ylab = "Area", #label y 
     type = "p", 
     pch = 20)
```

(4)(c) Present a horizontal, 
notched, 
colored boxplot of the areas calculated in (b). 
Title and label the axis.

```{r test3c}
?boxplot()
boxplot(p4b_area,
        horizontal =TRUE,
        notch = TRUE,
        col = c("cadetblue1", "black"),
        main = "Trees - Area",
        xlab = "Calculation - p4b_area", 
        ylab = "Trees Dataframe - # of values")
      
```

(4)(d) Demonstrate that the outlier revealed in the boxplot of Volume is not an extreme outlier.  
It is possible to do this with one line of code using *boxplot.stats()* 
or via 'manual' calculation and logicals. 

Identify the tree with the largest Volume and  
output on one line its row number and 
three measurements.

```{r test3d}

boxplot.stats(p4b_area, 
              coef = 3,
              do.conf = TRUE,
              do.out = TRUE)
#the output "$out" yielding numeric(0) means that the outlier is not extreme, with coefficient equal to 3

#leave columns blank to show all columns
#row 31, girth 20.6, height 87, volume 77
trees[which.max(p4b_area),] 


```

-----

##### Section 5: (12 points) The exponential distribution is an example of a right-skewed distribution with outliers.  This problem involves comparing it with a normal distribution which typically has very few outliers.  

5(a) Use *set.seed(124)* and 
*rexp()* with 
n = 100, 
rate = 5.5 to 
generate a random sample designated as y.  
Generate a second random sample designated as x with 
*set.seed(127)* and 
*rnorm()* using 
n = 100, 
mean = 0 and 
sd = 0.15.  

Generate a new object using *cbind(x, y)*. 
Do not output this object; instead, 
assign it to a new name. 
Pass this object to *apply()* and 
compute the inter-quartile range (IQR) for each column:  x and y. 
Use the function *IQR()* for this purpose. 
Round the results to four decimal places and present (this exercise shows the similarity of the IQR values.).

For information about *rexp()*, use 
*help(rexp)* or 
*?rexp()*.  
**Do not output x or y.**

```{r test5a}

set.seed(124)
y <- rexp(n = 100, rate = 5.5)

set.seed(127)
x <- rnorm(n = 100,  mean =  0, sd = 0.15)

xy <- cbind(x, y)

xy_IQR <- apply(xy, 2, IQR)

xy_IQR_round4 <- round(xy_IQR,4)


```

(5)(b) This item will illustrate the difference between a right-skewed distribution and 
a symmetric one.  
For base R plots, use *par(mfrow = c(2, 2))* to generate a display with four diagrams; 
*grid.arrange()* for ggplots. 
On the first row, for the normal results, present a histogram and 
a horizontal boxplot for x in color.  

For the exponential results, present a histogram and 
a horizontal boxplot for y in color.

```{r test5b}

#multi field row configuration 2x2 
par(mfrow = c(2, 2))

#normal results
hist(x,
     main = "Histogram - x",
     xlab = "rnorm(n = 100,  mean =  0, sd = 0.15)",
     ylab = "frequency",
     breaks = 10,
     col = "steelblue1",
     freq = FALSE)

boxplot(x,
        main = "Boxplot - x",
        horizontal = TRUE,
        col = "steelblue1")

#exponential results
hist(y,
     main = "Histogram - y",
     xlab = "rexp(n = 100, rate = 5.5)",
     ylab = "frequency",
     breaks = 10,
     col = "cadetblue1",
     freq = FALSE)

boxplot(y,
        main = "Boxplot - y",
        horizontal = TRUE,
        col = "cadetblue1")

```

(5)(c) QQ plots are useful for detecting the presence of heavy-tailed distributions. 
Present side-by-side QQ plots, 
one for each sample, 
using *qqnorm()* and *qqline()*.  
Add color and titles. 
In base R plots, "cex" can be used to control the size of the plotted data points and text. 
Lastly, determine if there are any extreme outliers in either sample. 
Remember:  extreme outliers are based on 3.0 x IQR in the box plot. 
R uses a default value of 1.5 x IQR to define outliers (not extreme) in both boxplot and boxplot stats.

```{r test5c}
par(mfcol = c(1,2))
#qqnorm x
qqnorm(x, 
       main = "QQ Norm - x",
       col = "steelblue1", 
       pch = 20)
#qqline x
qqline(x,
       col = "red",
       cex = 0.8)

#qqnorm y 
qqnorm(y, 
       main = "QQ Norm - y",
       col = "darkgreen",
       ylim = c(0,2), 
       pch = 20)
#qqline y 
qqline(y,
       col = "red",
       cex = 0.8)

#no outliers in x
boxplot.stats(x, coef = 3,
              do.conf = TRUE,
              do.out = TRUE)

#outlier in y 

boxplot.stats(y, coef = 3,
              do.conf = TRUE,
              do.out = TRUE)
##$out
##[1] 1.448679


```
